A Web-Based Data Visualization Platform for MATSim

### 2020-05-15 src/paper.tex
all content added and auto-references working

\documentclass[Afour,sageh,times]{sagej}

@@

\author{Billy Charlton and Janek Laudan\affilnum{1}}

@@

\keywords{MATSim, data visualization}

@@

MATSim is an open-source framework for implementing large-scale agent-based transport simulations (Horni, et. al, 2016). MATSim is widely used for transportation research in academic settings, and is gaining momentum as a tool ready for practice in real-world planning contexts.

There are many tools available for analyzing MATSim results, both
open-source and commercial. Typically, analysts can choose either
the free tool OTFVis (Strippgen, 2016) or the commercial software
Via (Rieser, 2016), both of which are desktop software packages requiring installation as well as a fair amount of technical acumen to operate. Alternatives to these tools include the more general-purpose desktop mapping software packages such as QGis and ArcGIS, or statistical software packages, again all of which require installation and expertise to use.

@@

\section{Requirements}
The research team at Technische Universität Berlin had several
"blank slate" discussions before any code was written: meaning,
if we could start at the very beginning and design something completely web-based and open, what would the bare minimum requirements be for it to be truly useful? The following requirements emerged from those discussions.

@@

\item[(i)] \textbf{HTML 5} improves and standardizes the "document model" of what constitutes a web page and how it is specified.
\item[(i)] \textbf{CSS 3} is a styling language that enables fine-grained layout and styling of individual elements on a page. CSS 3 defines in a consistent, standard way the details of elements including color, size, layout, and animation of page elements.
\item[(i)] \textbf{WebGL} provides in-browser support for the 3D-accelerated graphics capabilities of modern machines.
\item[(i)] \textbf{ECMAScript 6} is an updated (2015) version of the venerable Javascript scripting language that has been part of the web platform since the early 1990's. Recent versions of Javascript eliminate the more problematic aspects of the language and make it easier for developers to create bug-free, efficient code.
\item[(i)] \textbf{Web Workers} are a recent (2013) addition to the web platform that allow background thread processing for long-running tasks. Before Web Workers, there was no way to run truly multi-threaded code inside a browser.

@@

The software developed as part of this research is licensed entirely with the GNU General Public License v3, commonly known as the "GPL" (Free Software Foundation, 2007).

@@

Leaflet (leafletjs.com) is very popular and its application programming interface (API) is a bit simpler than that of Mapbox GL. Leaflet uses background map "tiles" at specific zoom levels, and layers data on top of those base maps. With small networks (we tested Cottbus, Germany, a small city of 100,000 inhabitants) Leaflet performed well, but medium-sized and large-sized networks with many elements visible at once suffered from noticeable performance degradation. This was problematic, as this was the simplest use-case envisioned.

Mapbox GL (mapbox.com) fared much better, apparently better-suited to displaying large datasets with many visible features simultaneously. In addition, Mapbox GL's use of 3D vector graphic mapping instead of preset tiles made for a much more pleasing user experience, with smooth animations between zoom levels and better background processing during page loads. For these reasons, Mapbox GL was chosen as the base map for the remaining geographic visualizations.

@@

After experimenting with several packages including D3, Raphael, Morris and others, the package Vega-Lite (Satyanarayan, 2016) exhibited many of the characteristics desired. Notably, Vega-lite follows a "grammar of graphics" declarative syntax, as popularized by Wilkinson (2005), and this grammar allows concise description of the meaningful components of a graphic. An added benefit is that graphs and charts can be downloaded in image format or in Vega source format, which is helpful for other researchers wanting to learn how to use the data format themselves.

@@

\item[(i)] Origin/destination flows between aggregate areas, so called "spider diagrams"
\item[(i)] Link flows, by time of day and mode
\item[(i)] Transit supply explorer, which displays all transit routes and allows the user to see which routes serve specific stops and links.
\item[(i)] Sankey diagrams, which can be used to depict changes/flows between between scenarios across multiple choices, such as shifts in mode between two scenarios (see Figure 1d)
\item[(i)] Emissions levels on a geographic hexagonal grid basis

@@

All code is available on the MATSim Github site, at https://github.com/matsim-org/viz.

@@

\subsection{References}
Please note that the files \textsf{SageH.bst} and \textsf{SageV.bst} are included with the class file
for those authors using \BibTeX.
The files work in a completely standard way, and you just need to uncomment one of the lines in the below example depending on what style you require:
\begin{verbatim}
%%Harvard (name/date)
%\bibliographystyle{SageH}
%%Vancouver (numbered)
%\bibliographystyle{SageV}
\bibliography{<YourBibfile.bib>}
\end{verbatim}
and remember to add the relevant option to the \verb+\documentclass[]{sagej}+ line as listed in Table~\ref{T1}.

%\section{Support for \textsf{\journalclass}}
%We offer on-line support to participating authors. Please contact
%us via e-mail at \dots
%
%We would welcome any feedback, positive or otherwise, on your
%experiences of using \textsf{\journalclass}.

\section{Copyright statement}
Please  be  aware that the use of  this \LaTeXe\ class file is
governed by the following conditions.

\subsection{Copyright}
Copyright \copyright\ \volumeyear\ SAGE Publications Ltd,
1 Oliver's Yard, 55 City Road, London, EC1Y~1SP, UK. All
rights reserved.

\subsection{Rules of use}
This class file is made available for use by authors who wish to
prepare an article for publication in a \textit{SAGE Publications} journal.
The user may not exploit any
part of the class file commercially.

This class file is provided on an \textit{as is}  basis, without
warranties of any kind, either express or implied, including but
not limited to warranties of title, or implied  warranties of
merchantablility or fitness for a particular purpose. There will
be no duty on the author[s] of the software or SAGE Publications Ltd
to correct any errors or defects in the software. Any
statutory  rights you may have remain unaffected by your
acceptance of these rules of use.

\begin{acks}
This class file was developed by Sunrise Setting Ltd,
Brixham, Devon, UK.\\
Website: \url{http://www.sunrise-setting.co.uk}
\end{acks}

\begin{thebibliography}{99}
\bibitem[Kopka and Daly(2003)]{R1}
Kopka~H and Daly~PW (2003) \textit{A Guide to \LaTeX}, 4th~edn.
Addison-Wesley.

\bibitem[Lamport(1994)]{R2}
Lamport~L (1994) \textit{\LaTeX: a Document Preparation System},
2nd~edn. Addison-Wesley.

\bibitem[Mittelbach and Goossens(2004)]{R3}
Mittelbach~F and Goossens~M (2004) \textit{The \LaTeX\ Companion},
2nd~edn. Addison-Wesley.

>>>>

\documentclass[Afour,sagev,times]{sagej}

@@

\author{Billy Charlton\affilnum{1} and Janek Laudan\affilnum{1}}

@@

\keywords{MATSim, data visualization, agent-based simulation}

@@

MATSim\cite{R1} is an open-source framework for implementing large-scale agent-based transport simulations. MATSim is widely used for transportation research in academic settings, and is gaining momentum as a tool ready for practice in real-world planning contexts.

There are many tools available for analyzing MATSim results, both
open-source and commercial. Typically, analysts can choose either
the free tool OTFVis\cite{R2} or the commercial software
Via \cite{R3}, both of which are desktop software packages requiring installation as well as a fair amount of technical acumen to operate. Alternatives to these tools include the more general-purpose desktop mapping software packages such as QGis and ArcGIS, or statistical software packages, again all of which require installation and expertise to use.

@@

\section{Requirements}
The research team at Technische Universit\"at Berlin had several
"blank slate" discussions before any code was written: meaning,
if we could start at the very beginning and design something completely web-based and open, what would the bare minimum requirements be for it to be truly useful? The following requirements emerged from those discussions.

@@

\item[(i)] \textbf{HTML 5} improves and standardizes the "document model" of what constitutes a web page and how it is specified.
\item[(i)] \textbf{CSS 3} is a styling language that enables fine-grained layout and styling of individual elements on a page. CSS 3 defines in a consistent, standard way the details of elements including color, size, layout, and animation of page elements.
\item[(ii)] \textbf{WebGL} provides in-browser support for the 3D-accelerated graphics capabilities of modern machines.
\item[(iv)] \textbf{ECMAScript 6} is an updated (2015) version of the venerable Javascript scripting language that has been part of the web platform since the early 1990's. Recent versions of Javascript eliminate the more problematic aspects of the language and make it easier for developers to create bug-free, efficient code.
\item[(v)] \textbf{Web Workers} are a recent (2013) addition to the web platform that allow baiun truly multi-threaded code inside a browser.

@@

The software developed as part of this research is licensed entirely with the GNU General Public License version 3, commonly known as the \textit{GPL v3}\cite{R4}.

@@

Leaflet (\url{leafletjs.com}) is very popular and its application programming interface (API) is a bit simpler than that of Mapbox GL. Leaflet uses background map "tiles" at specific zoom levels, and layers data on top of those base maps. With small networks (we tested Cottbus, Germany, a small city of 100,000 inhabitants) Leaflet performed well, but medium-sized and large-sized networks with many elements visible at once suffered from noticeable performance degradation. This was problematic, as this was the simplest use-case envisioned.

Mapbox GL (\url{mapbox.com}) fared much better, apparently better-suited to displaying large datasets with many visible features simultaneously. In addition, Mapbox GL's use of 3D vector graphic mapping instead of preset tiles made for a much more pleasing user experience, with smooth animations between zoom levels and better background processing during page loads. For these reasons, Mapbox GL was chosen as the base map for the remaining geographic visualizations.

@@

After experimenting with several packages including D3, Raphael, Morris and others, the package Vega-Lite\cite{R5} exhibited many of the characteristics desired. Notably, Vega-Lite follows a declarative syntax known as the "grammar of graphics", as popularized by Wilkinson\cite{R6}. This grammar allows concise description of the meaningful components of a graphic. An added benefit is that graphs and charts can be downloaded in image format or in Vega source format, which is helpful for other researchers wanting to learn how to use the data format.

@@

\item[(i)] Origin/destination flows between aggregate areas, so called "spider diagrams"
\item[(i)] Link flows, by time of day and mode
\item[(iii)] Transit supply explorer, which displays all transit routes and allows the user to see which routes serve specific stops and links.
\item[(iv)] Sankey diagrams, which can be used to depict changes/flows between between scenarios across multiple choices, such as shifts in mode between two scenarios (see Figure 1d)
\item[(v)] Emissions levels on a geographic hexagonal grid basis

@@

All code is available on the MATSim Github site, at \url{github.com/matsim-org/viz}.

@@

\subsection{References}
Please note that the files \textsf{SageH.bst} and \textsf{SageV.bst} are included with the class file
for those authors using \BibTeX.
The files work in a completely standard way, and you just need to uncomment one of the lines in the below example depending on what style you require:
\begin{verbatim}
%%Vancouver (numbered)
\bibliographystyle{SageV}
\bibliography{paper.bib}
\end{verbatim}
and remember to add the relevant option to the \verb+\documentclass[]{sagej}+ line as listed in Table~\ref{T1}.

\begin{acks}
This research project is part of the National Research Programme "Big Data" (NRP 75) of the Swiss National Science Foundation.\\
Website: \url{http://www.nrp75.ch/}
\end{acks}

\begin{thebibliography}{99}

\bibitem[Horni, A., Nagel, K., \& Axhausen, K. W. (Eds.). (2016)]{R2}
Horni, A., Nagel, K., \& Axhausen, K. W. (Eds.). (2016). The multi-agent transport simulation MATSim (p. 618). London: Ubiquity Press.

\bibitem[Strippgen, David (2016)]{R5}
Strippgen, David (2016) \textit{OTFVis: MATSim’s Open-Source Visualizer}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 225–234.

\bibitem[Rieser, Marcel (2016)]{R3}
Rieser, Marcel (2016) \textit{Senozon Via}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 219–224.

\bibitem[Free Software Soundation(2007)]{R1}
Free Software Soundation(2007) \textit{GNU General Public License}, Version 3.
URL: https://www.gnu.org/licenses/gpl.html

\bibitem[Satyanarayan, A., Moritz, D., Wongsuphasawat, K., and Heer, J. (2016)]{R4}
Satyanarayan, A., Moritz, D., Wongsuphasawat, K., and Heer, J. (2016) \textit{Vega-Lite: A Grammar of Interactive Graphics}. IEEE Transactions on Visualization and Computer Graphics, Volume 23, Issue 1. DOI: https://doi.org/10.1109/TVCG.2016.2599030

\bibitem[Wilkinson, Leland (2005)]{R6}
Wilkinson, Leland (2005) \textit{The Grammar of Graphics}, Second Edition. Springer Press, Chicago, USA. DOI: https://doi.org/10.1007/0-387-28695-0
